---
title: "Project proposal"
date: "Spring, 2023"
execute:
  warning: true
format: revealjs
---


## Project proposal steps

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(rvest)
courage = read_csv("data/courage.csv")
```

::: {style="font-size: 0.7em; text-align: left"}


1. Find a data set that satisfies [the guidelines](https://sta199-sp23-1.github.io/project-description.html).

2. Write about:

    - the source of data
    - when and how it was originally collected (by the curator, not necessarily how you found the data)
    - a brief description of the observations

3. Choose 1-2 research questions and write a hypothesis for each.

4. put your data in the data folder and push. `glimpse` the data

:::

## Demo: data in data folder

::: {style="font-size: 0.7em; text-align: left"}

**load data from your computer**

- upload button in R files pane

**read and write a `csv` to file from a URL**

- `dinosaur_dataset = read_csv("website-url-here")`
- `write_csv(dataset, "data/dinosaur_dataset.csv")`

- Note `read_csv` and `write_csv` are from the `readr` package and loaded automatically with `library(tidyverse)`

**read in an excel spreadsheet**

- `readxl::read_xlsx()`
- `readxl::read_xls()`

:::

## Ex: Introduction and Data

::: {style="font-size: 0.7em; text-align: left"}

Data set #1: NC Courage Homefield Advantage 
Our first data set comes from the [National Women's Soccer League (NSWL) Github](https://github.com/adror1/nwslR) and was sourced from [nwslsoccer.com](https://www.nwslsoccer.com/).

The dataset contains 78 observations (soccer games) played by the NC courage spanning three seasons: 2017, 2018, 2019. There are 10 variables in this dataset. Some of the variables we care about are `home_team`, `away_team`, and `result` (of the game).

:::

## Ex: Research question(s):

::: {style="font-size: 0.7em; text-align: left"}

Does NC Courage have a home-field advantage? We hypothesize that NC Courage is more likely to win on their home field than another team's field.

- To answer this question we will use information about the `home_team`, and the `result` of the game.

Does winning propagate winning? When NC Courage win a game, does it increase the probability of winning the very next game?

- To answer this question we will use information about the `result` of the game and the `game_number`.

:::

## Ex: Glimpse

```{r}
glimpse(courage)
```



## Remember

::: {style="font-size: 0.7em; text-align: left"}

- list data sets in your proposal **in your order of preference** 

- all work for the proposal should go into `proposal.qmd` within your project repo.

- render to pdf and submit to Gradescope, selecting all for the exercise "proposal"

- link your teammates in Gradescope when you submit

:::