---
title: "Welcome to Lab"
date: "January 11, 2023"
format: revealjs
---

## Introductions

```{r}
#| echo: false
#| warning: false
library(countdown)
```

-   Meet the TA!

-   Introduce yourself (icebreaker)

-   Follow along these slides on the course website (under slides): [sta199-sp23-1.github.io](https://sta199-sp23-1.github.io)

-   Bookmark this!
    It's the course website.

## What to expect in labs

-   Introduce lab assignment (5-10 minutes, longer today)
-   Work on the lab assignment (you can find it on the course website). You will work with others but your submission must be your own for the first several labs.
-   Typically you won't finish labs in-class and they will be due 1 week after they are released.

## Tips

-   Read **all** instructions on the lab.
-   One work strategy is to get through portions that you think will be most challenging (which initially might be the coding component) during lab when we can help you on the spot and leave the narrative writing until later.
-   Make use of [office hours](https://docs.google.com/spreadsheets/d/16gf9dQ_41V628vKvFdS95JBOUKYSjuxJBo7OzPye-SI/edit?usp=sharing). Before you need help!

## Beginnings

-   Create a GitHub account (unless you already have one) on <https://github.com/>
    -   one day you might want to show off your work, so choose a username you will be proud to show to a future employer.
-   tell us your username by taking [this survey](https://forms.gle/PAdD9QMkiXLCNRju5). This is **essential** to receive credit on future assignments.

```{r}
#| echo: false

countdown(minutes = 5, font_size = "1.5em",
          color_running_background = "#2774AE")
```

## Create your GitHub repo

Create your own copy of today's lab by clicking [here](https://classroom.github.com/a/FkaKJWlM). You may be prompted to login to GitHub. Your copy of of the lab is called your "repository", or "repo" for short.

## Demo: setting up

Now that your lab repo is created, let's setup git to work within RStudio.

To begin, open

- the lab instructions [here](https://sta199-sp23-1.github.io/labs/lab-0.html) and
- the RStudio containers [here](https://cmgr.oit.duke.edu/containers)
- your repo (that contains your starting files for the lab) [here](https://github.com/sta199-sp23-1)

Follow the instructions in the lab as I demo.

## Remember

No programming experience is required or expected to take this class. Come to lecture, lab and office hours. It's okay to be confused. Stick with it, you will get the hang of it!