---
title: "Lab 7 - Logistic regression"
categories: "Lab"
execute:
  eval: true
editor: 
  markdown: 
    wrap: 80
---

::: callout-important
This lab is due Monday, March 27 at 5:00pm.
:::

## Packages

You'll need the following packages for today's lab.

```{r}
#| warning: false
#| message: false
library(tidyverse)
library(tidymodels)
library(dsbox)
```

## Data

The data can be found in the **dsbox** package, and it's called `gss16`. Since
the dataset is distributed with the package, we don't need to load it
separately; it becomes available to us when we load the package.

If you would like to explicitly load the data into your environment so you can
view it, you can do so by running this code.

```{r}
#| warning: false
#| message: false
gss16 <- gss16
```

You can find out more about the dataset by inspecting its documentation, which
you can access by running `?gss16` in the Console or using the Help menu in
RStudio to search for `gss16`. You can also find this information
[here](https://rstudio-education.github.io/dsbox/reference/gss16.html).

# Exercises

## Exercise 1 - Data wrangling

::: callout-important
Remember: For each exercise, you should choose one person to type. All others
should contribute to the discussion, but only one person should type up the
answer, render the document, commit, and push to GitHub. All others should
**not** touch the document.
:::

a)  Create a new data frame called `gss16_advfront` that includes the variables
    `advfront`, `educ`, `polviews`, and `wrkstat`. Then, use the `drop_na()`
    function to remove rows that contain `NA`s from this new data frame.

b)  Transform the `advfront` variable such that it has two levels:
    `"Strongly agree"` and `"Agree"` should both be mapped to `"Agree"` and the
    remaining levels should all be relabeled `"Not agree"`. Make sure the
    resulting levels are in the following order: `"Agree"` and `"Not agree"`.

Hint: use the `factor()` function inside a `mutate()` statement to relabel the
original levels. Be sure to list the levels in order so that they are correctly
ordered after relabeling.

c)  Similarly to part b, combine the levels of the `polviews` variable such that
    levels that have the word "liberal" in them are lumped into a level called
    `"Liberal"` and those that have the word "conservative" in them are lumped
    into a level called `"Conservative"`. Make sure the levels are in the
    following order: `"Conservative"` , `"Moderate"`, and `"Liberal"`. Finally,
    `count()` how many times each new level appears in the `polviews` variable.

Hint: be careful if you manually type out the levels in the original `polviews`
variable to note that there are typos in two of the original levels "slightly
conservative" and "extremely conservative" are both misspelled, and so you will
need to match those misspellings in your call to `factor()`.

::: callout-important
After the team member working on Exercise 1 renders, commits, and pushes, all
other team members should pull. Then, choose a new team member to write the
answer to Exercise 2. (And so on for the remaining exercises.)
:::

## Exercise 2 - Train and test sets

Now, let's split the data into training and test sets so that we can evaluate
the models we're going to fit by how well they predict outcomes on data that
wasn't used to fit the models.

Specify a random seed of 1234 (i.e., include `set.seed(1234)` at the beginning
of your code chunk), and then split `gss16_advfront` randomly into a training
set `train_data` and a test set `test_data`. Do this so that the training set
contains 80% of the rows of the original data.

## Exercise 3 - Logistic Regression

a)  Using the training data, fit a logistic regression model that predicts
    `advfront` using `educ`. In particular, the model should predict the
    probability that `advfront` has value `"Not agree"`. Name this model
    `model1`. Report the tidy model output.

b)  Write out the fitted model equation in proper notation. State the meaning of
    any variables in the context of the data.

c)  Using your fitted model, report the estimated probability of agreeing with
    the following statement: Even if it brings no immediate benefits, scientific
    research that advances the frontiers of knowledge is necessary and should be
    supported by the federal government (`Agree` in advfront) if you have an
    education of 7 years.

## Exercise 4 - Another model

a)  Again using the training data, fit a new logistic regression model that adds
    the additional explanatory variable of `polviews`. Name this model `model2`.
    Report the tidy output.

b)  Now, report the estimated probability of agreeing with the following
    statement: Even if it brings no immediate benefits, scientific research that
    advances the frontiers of knowledge is necessary and should be supported by
    the federal government (`Agree` in advfront) if you have an education of 7
    years and are Conservative.

## Exercise 5 - Evaluating models with AIC

a)  Report the AIC values for each of `model1` and `model2`.

b)  Based on your results in part a, does it appear that including political
    views in addition to years of education is useful for modeling whether
    employees agree with the statement "Even if it brings no immediate benefits,
    scientific research that advances the frontiers of knowledge is necessary
    and should be supported by the federal government"? Explain.

## Exercise 6 - Evaluating models using test data

a)  For each of `model1` and `model2`, report the number of false positive and
    false negatives when making predictions on the `test_data` with a decision
    boundary of 0.5.

b)  Do these results provide much information about which model you would prefer
    for a prediction task? If so, which model would you choose?

c)  Do you think a decision boundary of 0.5 makes sense here or would you adjust
    it?

The ROC curve provides a way to compare predictive performance of binary
classifiers across the full range of decision boundaries. Notes about the ROC
curve can be found here <https://sta199-s23-2.github.io/ae-sa/ae-16-A.html>.

# Submission

::: callout-warning
Before you wrap up the assignment, make sure all documents are updated on your
GitHub repo. We will be checking these to make sure you have been practicing how
to render and push changes.

You must turn in a PDF file to the Gradescope page by the submission deadline to
be considered "on time". Only one team member should submit to Gradescope, but
they should add all other team members to the submission.

Make sure your data are tidy! That is, your code should not be running off the
pages and spaced properly. See: <https://style.tidyverse.org/ggplot2.html>.
:::

To submit your assignment:

-   Go to <http://www.gradescope.com> and click *Log in* in the top right
    corner.
-   Click *School Credentials* $\rightarrow$ *Duke NetID* and log in using your
    NetID credentials.
-   Click on your *STA 199* course.
-   Click on the assignment, and you'll be prompted to submit it.
-   Mark all the pages associated with exercise. All the pages of your lab
    should be associated with at least one question (i.e., should be "checked").
    *If you do not do this, you will be subject to lose points on the
    assignment.*
-   *Select all* pages of your .pdf submission to be associated with the
    *"Workflow & formatting"* question.

# Grading

| Component             | Points |
|:----------------------|:-------|
| Ex 1                  | 8      |
| Ex 2                  | 3      |
| Ex 3                  | 10     |
| Ex 4                  | 5      |
| Ex 5                  | 5      |
| Ex 6                  | 14     |
| Workflow & formatting | 5      |
| **Total**             | 50     |

::: callout-note
The "Workflow & formatting" grade is to assess the reproducible workflow. This
includes:

-   linking all pages appropriately on Gradescope

-   putting your team and member names in the YAML at the top of the document

-   committing the submitted version of your `.qmd` to GitHub

-   Are you under the 80 character code limit? (You shouldn't have to scroll to
    see all your code). Pipes `%>%`, `|>` and ggplot layers `+` should be
    followed by a new line

-   You should be consistent with stylistic choices, e.g. only use 1 of `=` vs
    `<-` and `%>%` vs `|>`

-   All binary operators should be surrounded by space. For example `x + y` is
    appropriate. `x+y` is not.
:::
